Left-linear rules in adhesive categories

### 2024-02-11 icalp/fusione.tex
cose

\documentclass[a4paper,UKenglish,cleveref,pdftex, thm-restate,numberwithinsect]{lipics}
\newif\ifreport
\reporttrue
%\reportfalse

% full or conference version
\newif\iffull
\fulltrue
%\fullfalse

@@

%\tableofcontents

@@

\subsection{$\mathcal{M}$-adhesivity}
The key property that $\mathcal{M}$-adhesive categories enjoy is given
by the so-called \emph{Van Kampen condition}
\cite{brown1997van,johnstone2007quasitoposes,lack2005adhesive}. We
will recall it and examine some of its consequences. First of all we
need to recall some terminology and facts regarding subclasses of
$\mor(\X)$.

@@

%<<<<<<< HEAD
%\todo{P: qui probabilmente si puo' compattare tenendo solo cio' che si usa}
%\todo{Si usa tutto per la classe B+ [DC]}
%$\mathcal{M}$-adhesivity is well-behaved with respect to  the comma construction \cite{mac2013categories}, as shown by the following theorem.
%\begin{theorem}[\cite{ehrig2006fundamentals,lack2005adhesive}]\label{lem:comma}
  %Let $\A$ and $\B$ be respectively an $\mathcal{M}$-adhesive and an
  %$\mathcal{M}'$-adhesive category. Let also $L:\A\rightarrow \C$ be a
  %functor that preserves $\mathcal{M}$-pushouts, and
  %$R:\B\rightarrow \C$ be a functor which preserves pullbacks.Then
 % $\comma{L}{R}$ is $\cma{M}{M'}$-adhesive, where
  %\[
    %\cma{M}{M}':=\{(h,k)\in \mathcal{A}(\comma{L}{R}) \mid h\in
    %\mathcal{M}, k\in \mathcal{M}'\}\]
%\end{theorem}
$\mathcal{M}$-adhesivity is well-behaved with respect to  the construction of slice and functor caregories, \cite{mac2013categories}, as shown by the following theorems \cite{ehrig2006fundamentals,lack2005adhesive}.
   --ò-à

@@

\subsection{DPO rewriting systems derivations}

@@

  We will call the number $n+1$ the \emph{length} of the derivation,
  denoted by $\lgh(\der{D})$. We will also say that an empty
  derivation has length $0$.\todo{P: used}

  Moreover, if every $\dder{D}_i$ applies the rule $\rho_i\in R$, then
  we can define an associated sequence of rules as $r(\der{D})$ as
  $\{\rho_i\}_{i=0}^n$.\todo{P: Used?}
\end{definition}

@@

Given DPO rewriting system $(\X, \R)$, we have already noted in
\Cref{rem:func} that a derivation $\der{D}$ determines a diagram
$\Delta(\der{D})$ in $\X$. We can then wonder if such a diagram has a
colimit. Clearly if $\der{D}$ is the empty derivation $G$ then a
colimit for $\Delta(\der{D})$ is simply the object $G$. More
generally, we have the following result.

\begin{lemma}\label{lem:colim}
  Let $\X$ be an $\mathcal{M}$-adhesive category and $(\X, \R)$ a
  left-linear DPO rewriting system over it. The following properties
  hold true.
  \begin{enumerate}
  \item  If $\der{D}$ is a derivation from $G$ to $H$, then the diagram $\Delta(\der{D})$ has a colimit $(\tpro{D}, \{\iota_X\}_{X\in \Delta(\der{D})})$ such that $\iota_H$ belongs to $\mathcal{M}$.
  \item Let $\der{D}$ be the concatenation $\der{D}_1\cdot \der{D}_2$ of two derivations $\der{D}_1=\{\dder{D}_{1,i}\}_{i=0}^{n_1}$ between $G$ and $H$ and $\der{D}_2=\{\dder{D}_{2,j}\}_{j=0}^{n_2}$ between $H$ and $T$,  then the colimiting cocone $(\tpro{D}, \{\iota_X\}_{X\in \Delta(\der{D})})$ exists too and there is a pushout square
    \[\xymatrix{H\ar[r]^-{\iota_{2, H}} \ar@{>->}[d]_-{\iota_{1, H}} & \tproi{D}{2} \ar[d]^{p_2}\\  \tproi{D}{1} \ar[r]_{p_1}& \tpro{D}}\]
    where $(\tproi{D}{1}, \{\iota_{1, X}\}_{X\in \Delta(\der{D}_1)})$ and $(\tproi{D}{2}, \{\iota_{2, X}\}_{X\in \Delta(\der{D}_2)})$ are the colimiting cocone for $\Delta(\der{D}_1)$ and $\Delta(\der{D}_2)$, respectively.
  \end{enumerate}
\end{lemma}
\begin{remark}\label{rem:cof}
  Let $I:\Deltamin(\dder{D})\to \Delta(\dder{D})$ be the inclusion functor. It is immediate to see that such functor is \emph{final} \cite{mac2013categories}. This means that for every functor $F\colon \Delta(\dder{D})\to \Y$ we have:
  \begin{enumerate}
  \item if  $(C, \{c_X\}_{X\in \Deltamin(\dder{D})})$ is colimiting for $F\circ I$, then there exists a colimiting cocone $(D, \{d_X\}_{X\in \Delta(\dder{D})})$ for $F$;
  \item $(C, \{c_X\}_{X\in \Deltamin(\dder{D})})$ and $(D, \{d_X\}_{X\in \Delta(\dder{D})})$ are colimiting for, repsectively, $F\circ I$ and $F$, then the canonical arrow $\phi\colon C\to D$ induced by $(D, \{d_X\}_{X\in \Deltamin(\dder{D})})$ is an isomorphism.
  \end{enumerate}
\end{remark}

    \smallskip \noinden
    t$\lgh(\dder{D})=1$. Suppose that $\dder{D}$
    has as its single component the derivation
    \[\xymatrix{L \ar[d]_{m}& K \ar[d]^{k}\ar@{>->}[l]_{l} \ar[r]^{r} & R \ar[d]^{h} \\G& \ar@{>->}[l]^{f} D \ar[r]_{g}& H  \\}\]
    The arrow $f$ is  the pushout of $l$ and so it is in in $\mathcal{M}$. We can thus consider the $\mathcal{M}$-pushout square
    \[\xymatrix{D \ar@{>->}[d]_{f} \ar[r]^{g} & H \ar@{>->}[d]^{p} \\G \ar[r]_{q}& P }\]
    Since $p\in \mathcal{M}$, the thesis follows immediately from \Cref{rem:cof}.

    Let us now show that the square above is a pushout. Take two arrows $a\colon \tproi{D}{1}\to C$, $b\colon \tproi{D}{2}\to C$ such that
    \[a\circ \iota_{1, H}=b\circ \iota_{2, H}\]
    We can use the previous equality to define a cocone $(C, \{c_X\}_{X\in \Delta(\der{D})})$ putting:
    \[c_X:=\begin{cases}
        a\circ \iota_{1, X} & X\in \Delta(\der{D}_1) \\
        b\circ \iota_{2, X} & X\in \Delta(\der{D}_2)
      \end{cases}\]
    From this, we can deduce at once the existence of a unique $c\colon \tpro{D}\to C$ such that
    \[c\circ \iota_X = c_X\]
    By construction, for every $X\in  \Delta(\der{D}_1)$ and $Y\in  \Delta(\der{D}_2)$ we have
    \[\begin{split}
        c\circ p_1 \circ \iota_{1,X}&=c\circ \iota_{X}\\&=c_X \\&=a\circ \iota_{1,X}\\&=a\circ p_1\circ \iota_{1,X}
      \end{split}\qquad \begin{split}
        c\circ p_2 \circ \iota_{2,Y}&=c\circ \iota_{Y}\\&=c_Y \\&=b\circ \iota_{2,Y}\\&=b\circ p_2\circ \iota_{2,Y}
      \end{split}\]
    Therefore
    \[c\circ p_1=a \qquad c\circ p_2 = b\]

    For uniqueness, suppose that $c'\colon \tpro{D}\to C$ is such that
    \[c'\circ p_1=a \qquad c'\circ p_2 = b\]
    Then, for every $X\in \Delta(\der{D})$ we have
    \begin{align*}
      c'\circ \iota_X & = \begin{cases}
        c'\circ p_1\circ \iota_{1, X} & X\in \Delta(\der{D}_1) \\
        c'\circ p_2\circ \iota_{2, X} & X\in \Delta(\der{D}_2)
      \end{cases} \\&=\begin{cases}
        a\circ \iota_{1, X} & X\in \Delta(\der{D}_1) \\
        b\circ \iota_{2, X} & X\in \Delta(\der{D}_2)
      \end{cases}\\&=c_X\\&=c\circ \iota_X
    \end{align*}
    showing that $c'=c$ as wanted.	 \qedhere
  \end{enumerate}
\end{proof}

\begin{corollary}
  \label{cor:colim}
  Let $\der{D}=\{\dder{D}_i\}_{i=0}^n$ a derivation of length $n+1$
  and fix an index $j\in[0,n]$. Define
  \[\der{D}^j_1:=\{\dder{D}_i\}_{i=0}^{j-1} \qquad
    \der{D}^j_2=\{\dder{D}_j\} \qquad
    \der{D}^j_3:=\{\dder{D}_i\}_{i=j+1}^n
  \]
  with the convention that
  $\der{D}^0_1$ and $\der{D}^n_3$ are the empty derivation on,
  respectively, $G_0$ and $G_n$. Then the square below is a pushout
  and a pullback
  \[\xymatrix@C=30pt{D_{j} \ar[ddrr]^{\iota_{D_j}}
      \ar[r]^{g_j}\ar@{>->}[d]_{f_j}& G_{j+1}
      \ar[ddr]^{\iota_{G_{j+1}}}
      \ar[r]^-{\iota_{3, G_{j+1}}} & \lpro \der{D}^j_3\rpro \ar[dd]^{p_2} \\
      G_j\ar[drr]_{\iota_{G_{j}}} \ar@{>->}[d]_{\iota_{1,G_j}}\\ \lpro
      \der{D}^j_1 \rpro \ar[rr]_{p_1} &&\tpro{D} }
  \]
  Where the two arrows $p_1\colon \lpro\der{D}^j_1 \rpro\to \tpro{D}$,
  $p_2\colon \lpro \der{D}^j_3\rpro \to \tpro{D}$ are induced by the
  cocones $(\tpro{D}, \{\iota_{X}\}_{X\in \Delta(\der{D}^j_1)})$ and
  $(\tpro{D},\{\iota_{X}\}_{X\in \Delta(\der{D}^j_3)})$, respectively.
\end{corollary}
\begin{remark}
  If $\der{D}$ is empty then $\der{D}^j_1, \der{D}^j_2$ and
  $\der{D}^j_3$ are empty too.
\end{remark}

\begin{remark}
  \label{rem:zero1}
  In particular, considering $j=0$ or $j=n$, we get that the following
  two squares are $\mathcal{M}$-pushouts and, thus, pullbacks.
  \[
    \xymatrix{D_0 \ar[r]^-{\iota_{3, D_0}} \ar@{>->}[d]_{f_0}& \lpro
      \der{D}^0_3 \rpro \ar@{>->}[d]^{p_2} &D_n
      \ar@{>->}[d]_{\iota_{1, D_n}}\ar[r]^{g_n}& G_n
      \ar@{>->}[d]^{\iota_{G_n}}\\ G \ar[r]_{\iota_G}& \tpro{D} &
      \lpro \der{D}^n_1 \rpro \ar[r]_{p_1} & \tpro{D}}
  \]
\end{remark}

\begin{proof}
  Consider the diagram
  \[
    \xymatrix@C=34pt{T \ar@/_.6cm/[dd]_{a}\ar@{.>}[d]^{c}\ar[rr]^{b}&&
      G_k \ar[d]_{\iota_{3, G_k}}
      \ar@/^.6cm/[ddd]^{\iota_{G_k}}\\D_{j} \ar[r]^{g_j}
      \ar[ddrr]^{\iota_{D_j}}\ar@{>->}[d]^{f_j}& G_{j+1}
      \ar[r]^-{\iota_{3,G_{j+1}}} \ar[ddr]^{\iota_{G_{j+1}}} & \lpro
      \der{D}^j_3\rpro \ar[dd]_(.6){p_2} \\ G_j
      \ar@{>->}[d]_{\iota_{1,G_j}} \ar[drr]_{\iota_{G_j}}\\ \lpro
      \der{D}^j_1 \rpro \ar[rr]_{p_1} &&\tpro{D} }
  \]
  Thanks to \Cref{cor:colim} we know that the bottom right rectangle
  in the diagram above is a pullback and the thesis follows at once.
\end{proof}

@@

\subsection{Sequentially independent and switchable derivations }

@@

\begin{definition}[Switch]
  \label{def:switch}
  Let with $\X$ an $\mathcal{M}$-adhesive category 
  and $(\X, \R)$ a left-linear DPO rewriting system. 
  Let also
  $\der{D}=\{\dder{D}_i\}_{i=0}^1$ be a derivation of length $2$ made
  by two sequentially independent derivations $\dder{D}_0$,
  $\dder{D}_1$, with independence pair $(i_0, i_1)$, as in the diagram
  on the left below. A \emph{switch} of $\der{D}$ along $(i_0,i_1)$ is
  a derivation $\der{E}=\{\dder{E}_i\}_{i=0}^1$, using the same rules
  in reverse order, as in the diagram on the right below, 
  \[
    \def\objectstyle{\scriptstyle}\def\labelstyle{\scriptstyle}
    \xymatrix@C=1.1mm@R=8mm{
      {L_0} \ar[d]_{m_0}
      & & {K_0} \ar@{>->}[ll]_{l_0} \ar[rr]^{r_0} \ar[d]_{k_0}
      & & {R_0} \ar[drr]_{h_0}  \ar@(r,ul)@{.>}^(.75){i_0}[drrrrrr]
      & & & & 
      % 
      {L_1}\ar[dll]^{m_1} \ar@(l,ur)@{.>}_(.75){i_1}[dllllll]
      & & {K_1} \ar@{>->}[ll]_{l_1} \ar[rr]^{r_1} \ar[d]^{k_1}
      & & {R_1} \ar[d]^{h_1} \\
      % 
      % 
      {G}
      & & {D_0} \ar@{>->}[ll]^{f_0} \ar[rrrr]_{g_0}
      & & & & {H} & &
      & &  {D_1} \ar@{>->}[llll]^{f_1} \ar[rr]_{g_1}
      & & {H}
    }
    % 
    \xymatrix@C=1.1mm@R=8mm{
      {L_1} \ar[d]_{m_1'}
      & & {K_1} \ar@{>->}[ll]_{l_1} \ar[rr]^{r_1} \ar[d]_{k_1'}
      & & {R_1} \ar[drr]_{h_1'}  \ar@(r,ul)@{.>}^(.75){i_1'}[drrrrrr]
      & & & & 
      % 
      {L_0}\ar[dll]^{m_0'} \ar@(l,ur)@{.>}_(.75){i_0'}[dllllll]
      & & {K_0} \ar@{>->}[ll]_{l_0} \ar[rr]^{r_0} \ar[d]^{k_0'}
      & & {R_0} \ar[d]^{h_0'} \\
      % 
      % 
      {G}
      & & {D_1'} \ar@{>->}[ll]^{f_1'} \ar[rrrr]_{g_1'}
      & & & & {H'} & &
      & &  {D_0'} \ar@{>->}[llll]^{f_0'} \ar[rr]_{g_0'}
      & & {H'}  }
    %
  \]
  such that there is an independence pair $(i_1', i_0')$ between
  $\dder{E}_0$ and $\dder{E}_1$ and 
  \begin{center}   
    $m_0=f_0' \circ i_0'$
    \qquad $h_1=g_1' \circ i_1'$
    \qquad $m_0'= f_0 \circ i_1$
    \qquad $h_1'= g_{1}\circ i_0$.
  \end{center}

@@

As already mentioned, when viewing derivations as concurrent computations, 
for linear graph rewriting systems it is canonical to identify  derivations which are equal 
``up to switching''. The same notion can be given for left-linear rewriting systems relying on notion of switch.

@@

\subsection{Relating sequential indepedence and switchability}

@@

When working with left-linear rewriting systems, sequential independence does not imply 
switchability (while the converse implication clearly holds), as shown by the example below.

@@

\begin{example}
  \label{ex:diff1}
  Recall that, given two posets $(P, \leq)$ and $(Q, \leq)$, we can
  construct the poset $(P, \leq)\oplus(Q,\leq)$ which has as
  underlying set the disjoint union $P+Q$ and in which every element
  of $Q$ is greater than every element of $P$, while elements of $P$
  and $Q$ are compared using the original orders.

  Take now the set $S:=\{a,b,c\}$ and endow it with the smallest order
  such that $a\leq b$ and $a\leq c$.  Take also the set $\mathbb{N}$
  ordered by $\geq$, so that $0$ is its maximum.  Let $\X$ be the
  category associated to the order
  $(S, \leq)\oplus (\mathbb{N}, \geq)$, which by \Cref{rem:iso} is
  $\mathsf{Iso}$-adhesive. Notice that in this category the arrows
  $(a,b)\colon a\to b$ $(a,c)\colon a\to c$ do not have a pushout
  because $b$ and $c$ do not have a supremum.  Define $\R$ as the set
  containing the following two rules
  \[\xymatrix{a & a \ar[r]^{(a, 0)} \ar[l]_{\id{a}} & 0 & a & a
      \ar[r]^{(a, b)} \ar[l]_{\id{a}} & b }\]
  We can then consider the following derivation
  $\der{D}=\{\dder{D}_i\}_{i=0}^1$ of length $2$, made by two
  sequential independent direct derivations
  \[
    \xymatrix@C=30pt{a \ar[d]_{m_0}&& a \ar[d]_{k_0}\ar[ll]_{\id{a}}
      \ar[r]^{(a,0)} & 0 \ar@/^.35cm/[drrr]|(.315)\hole_(.45){\id{0}}
      \ar[dr]|(.3)\hole_{\id{0}} && a \ar@/_.35cm/[dlll]^(.45){(a,c)}
      \ar[dl]|(.3)\hole^{(a,0)}& a \ar[d]^{(a,0)}\ar@{>->}[l]_{\id{a}}
      \ar[rr]^{(a,b)} && b \ar[d]^{(b,0)}\\c &&
      \ar@{>->}[ll]^{(\id{c})} c \ar[rr]_{(c,0)}&& 0 &&
      \ar@{>->}[ll]^{\id{0}} 0 \ar[rr]_{\id{0}}&& 0}\]

  \noindent
  \parbox{10cm}{ \hspace{15pt}On the other hand, the rule
    applied by $\dder{D}_1$ cannot be applied to $c$. indeed the
    unique morphism $a\to c$ is $(a,c)$, thus we get a diagram
    aside. But $b$ and $c$ do not have a supremum in the poset
    underlying $\X$, thus we do not get a direct derivation from
    $c$.}  \parbox{3cm}{
    \vspace{-.3cm}$\xymatrix{a \ar[d]_{(a,c)}& a
      \ar[d]_{(a,c)}\ar@{>->}[l]_{\id{a}} \ar[r]^{(a,b)} & b \\c
      & c \ar[l]^{\id{c}} }$}  
\end{example}

@@

  The category $\cat{Graph}$ is tame by the result in ...  Thus,
  sequential independence ensure the existence of a switch $\der{E}$,
  depicted in Fig.~\ref{fi:derE}.
  \begin{figure}
    \input{Figure/derE.tex}
    \caption{The derivation $\der{E}$.}
    \label{fi:derE}
  \end{figure}
\end{example}

@@

\begin{definition}[Good independence pair]
  \label{def:filler}
  Let with $\X$ an $\mathcal{M}$-adhesive category 
  and $(\X, \R)$ a left-linear DPO rewriting system. 
  Let also $(i_0, i_1)$ be an independence pair between two direct
  derivations $\dder{D}_0$, $\dder{D}_1$
  as in the solid part of the diagram below
  \[
    \xymatrix@C=15pt{L_0 \ar[d]_{m_0}&& K_0
      \ar[d]_{k_0}\ar@{>->}[ll]_{l_0} \ar[r]^{r_0} \ar@/^.20cm/[ddrr]|(.3)\hole|(.55)\hole^(.7){u_0} & R_0
      \ar@/^.35cm/[drrr]|(.3)\hole_(.4){i_0} \ar[dr]|(.3)\hole_{h_0}
      && L_1 \ar@/_.35cm/[dlll]^(.4){i_1} \ar[dl]|(.3)\hole^{m_1}& K_1
      \ar[d]^{k_1}\ar@{>->}[l]_{l_1} \ar[rr]^{r_1} \ar@/_.20cm/[ddll]|(.3)\hole|(.55)\hole_(.7){u_1} && R_1 \ar[d]^{h_1} \\
      %
      G_0 &&
      \ar@{>->}[ll]^{f_0} D_0 \ar[rr]_{g_0}&& G_1 &&
      \ar@{>->}[ll]^{f_1} D_1 \ar[rr]_{g_1}&& G_2\\
      %
      & & & & P \ar@/^.3cm/@{>->}[ull]^{p_0} \ar@/_.3cm/[urr]_{p_1}
    }
  \]

  Consider the pullback of $g_0$ and $f_1$, getting $p_i : P \to D_i$
  ($i \in \{0,1\}$) and the arrows $u_0\colon K_0\to P$ and
  $u_1\colon K_1\to P$ given by \Cref{prop:tec} .
  %
  We will say that $(i_0, i_1)$ is a \emph{good independence pair} if
  the first and second square below are pushouts and if the third
  square exists and it is a pushout too
  \[
    \xymatrix{
      %
      K_0 \ar[r]^{r_0} \ar[d]_{u_0}& R_0 \ar[d]^{i_0} & K_1
      \ar@{>->}[r]^{l_1} \ar[d]_{u_1}& L_1 \ar[d]^{i_1}&K_1
      \ar[r]^{r_1} \ar[d]_{u_1}& R_1\ar@{.>}[d]^{j_0} \\
      %
      P \ar[r]_{p_1} & D_1 &P \ar[r]_{p_0} & D_0&P \ar@{.>}[r]_{q_1} &
      Q_1}
  \]
\end{definition}

@@

@@

@@

>>>>

\documentclass[a4paper,UKenglish,cleveref,pdftex, thm-restate,numberwithinsect,anonymous]{lipics}
\newif\ifreport
\reporttrue
%\reportfalse

% full includes some additional material
\newif\iffull
\fulltrue
%\fullfalse

@@

\tableofcontents

@@

\subsection{$\mathcal{M}$-adhesivity}\label{subsec:ade}
The key property that $\mathcal{M}$-adhesive categories enjoy is given
by the so-called \emph{Van Kampen condition}
\cite{brown1997van,johnstone2007quasitoposes,lack2005adhesive}. We
will recall it and examine some of its consequences. First of all we
need to recall some terminology and facts regarding subclasses of
$\mor(\X)$.

@@

%\todo{P: qui probabilmente si puo' compattare tenendo solo cio' che si usa}
%\todo{Si usa tutto per la classe B+ [DC]}
%$\mathcal{M}$-adhesivity is well-behaved with respect to  the comma construction \cite{mac2013categories}, as shown by the following theorem.
%\begin{theorem}[\cite{ehrig2006fundamentals,lack2005adhesive}]\label{lem:comma}
  %Let $\A$ and $\B$ be respectively an $\mathcal{M}$-adhesive and an
  %$\mathcal{M}'$-adhesive category. Let also $L:\A\rightarrow \C$ be a
  %functor that preserves $\mathcal{M}$-pushouts, and
  %$R:\B\rightarrow \C$ be a functor which preserves pullbacks.Then
 % $\comma{L}{R}$ is $\cma{M}{M'}$-adhesive, where
  %\[
    %\cma{M}{M}':=\{(h,k)\in \mathcal{A}(\comma{L}{R}) \mid h\in
    %\mathcal{M}, k\in \mathcal{M}'\}\]
%\end{theorem}
$\mathcal{M}$-adhesivity is well-behaved with respect to  the construction of slice and functor caregories, \cite{mac2013categories}, as shown by the following theorems \cite{ehrig2006fundamentals,lack2005adhesive}.
   --ò-à

@@

\subsection{DPO rewriting systems derivations}\label{subsec:DPO}

@@

  We will call the number $n+1$ the \emph{length} of the derivation,
  denoted by $\lgh(\der{D})$. We will also say that an empty
  derivation has length $0$.

  Moreover, if every $\dder{D}_i$ applies the rule $\rho_i\in R$, then
  we can define an associated sequence of rules as $r(\der{D})$ as
  $\{\rho_i\}_{i=0}^n$.
\end{definition}

@@

Given DPO rewriting system $(\X, \R)$,  a derivation $\der{D}$ determines a diagram
$\Delta(\der{D})$ in $\X$. Using $\mathcal{M}$-adhesivity, it can be shown  that such diagram has a colimit $(\tpro{D}, \{\iota_X\}_{X\in \Delta(\der{D})})$ and that, if $(\X, \R)$ is linear, then every coprojection is in $\mathcal{M}$ ( see \Cref{lem:colim} for details and a proof of these statements).

@@

\subsection{Sequentially independent and switchable derivations }\label{subsec:switch}

@@

\begin{definition}[Switch]
  \label{def:switch}
  Let with $\X$ an $\mathcal{M}$-adhesive category 
  and $(\X, \R)$ a left-linear DPO rewriting system. 
  Let also
  $\der{D}=\{\dder{D}_i\}_{i=0}^1$ be a derivation of length $2$ made
  by two sequentially independent derivations $\dder{D}_0$,
  $\dder{D}_1$, with independence pair $(i_0, i_1)$, as in the diagram
  on the left below. A \emph{switch} of $\der{D}$ along $(i_0,i_1)$ is
  a derivation $\der{E}=\{\dder{E}_i\}_{i=0}^1$, between the same objects and using the same rules
  in reverse order, as in the diagram on the right below, 
  \[
    \def\objectstyle{\scriptstyle}\def\labelstyle{\scriptstyle}
    \xymatrix@C=1.1mm@R=8mm{
      {L_0} \ar[d]_{m_0}
      & & {K_0} \ar@{>->}[ll]_{l_0} \ar[rr]^{r_0} \ar[d]_{k_0}
      & & {R_0} \ar[drr]|(.35)\hole_(.6){h_0}  \ar@/^.22cm/[drrrrrr]|(.3)\hole^(.75){i_0}
      & & & & 
      % 
      {L_1}\ar[dll]|(.35)\hole^(.6){m_1} \ar@/_.22cm/_(.75){i_1}[dllllll]
      & & {K_1} \ar@{>->}[ll]_{l_1} \ar[rr]^{r_1} \ar[d]^{k_1}
      & & {R_1} \ar[d]^{h_1} \\
      % 
      % 
      {G_0}
      & & {D_0} \ar@{>->}[ll]^{f_0} \ar[rrrr]_{g_0}
      & & & & {G_1} & &
      & &  {D_1} \ar@{>->}[llll]^{f_1} \ar[rr]_{g_1}
      & & {G_0}
    }
    % 
    \xymatrix@C=1.1mm@R=8mm{
      {L_1} \ar[d]_{m_0'}
      & & {K_1} \ar@{>->}[ll]_{l_1} \ar[rr]^{r_1} \ar[d]_{k_0'}
      & & {R_1} \ar[drr]|(.35)\hole_(.6){h_0'}  \ar@/^.22cm/@{.>}^(.75){i_0'}|(.3)\hole[drrrrrr]
      & & & & 
      % 
      {L_0}\ar[dll]|(.35)\hole^(.6){m_1'} \ar@/_.22cm/@{.>}_(.75){i_1'}[dllllll] 
      & & {K_0} \ar@{>->}[ll]_{l_0} \ar[rr]^{r_0} \ar[d]^{k_1'}
      & & {R_0} \ar[d]^{h_1'} \\
      % 
      % 
      {G_0}
      & & {D_1'} \ar@{>->}[ll]^{f_0'} \ar[rrrr]_{g_0'}
      & & & & {G'_1} & &
      & &  {D_0'} \ar@{>->}[llll]^{f_1'} \ar[rr]_{g_1'}
      & & {G_2}  }
    %
  \]
  such that there is an independence pair $(i_0', i_1')$ between
  $\dder{E}_0$ and $\dder{E}_1$ and 
  \begin{center}   
    $m_0=f_0' \circ i_1'$
    \qquad $h_1=g_1' \circ i_0'$
    \qquad $m_0'= f_0 \circ i_1$
    \qquad $h_1'= g_{1}\circ i_0$.
  \end{center}

@@

It can be shown that switches along the same independence pair are unique up to abstraction
equivalence (see Proposition~\ref{thm:switch_uni}).

  A derivation $\der{D}$ consisting of three steps $\dder{D}_0$,
  $\dder{D}_1$, and $\dder{D}_2$, each applying the corresponding rule
  $\rho_i$ ,is depicted in in Fig.~\ref{fi:derD}. The second and third
  steps $\dder{D}_1$ and $\dder{D}_2$ are clearly sequential
  independent, via an independence pairs that map nodes consistently
  with their numbering.\todo{P: having names for the graphs would
    probably help}

  The independence pair can be used to construct a switch $\der{E}$, as
  depicted in Fig.~\ref{fi:derE}.
  \begin{figure}
    \input{Figure/derE.tex}
    \caption{The derivation $\der{E}$.}
    \label{fi:derE}
  \end{figure}
\end{example}

As already mentioned, for linear rewriting systems, when viewing
derivations as concurrent computations, it is canonical to identify
derivations which are equal ``up to switching''. The same notion can
be given for left-linear rewriting systems relying on notion of
switch.

@@

\subsection{Church-Rosser: Relating sequential indepedence and switchability}\label{subsec:CR}

@@

Unfortunately, when working with left-linear rewriting systems,
sequential independence does not imply switchability (while the
converse implication clearly holds), as shown by the example below.

@@

\begin{example}
  \label{ex:diff1}
  Given two posets $(P, \leq)$ and $(Q, \leq)$, define
  $(P, \leq)\oplus(Q,\leq)$ as the poset which has as underlying set
  the disjoint union $P+Q$ and in which every element of $Q$ is
  greater than every element of $P$, while elements of $P$ and $Q$ are
  compared using the original orders.

  Take now the set $S:=\{a,b,c\}$ and endow it with the smallest order
  such that $a\leq b$ and $a\leq c$.  Take also the set $\mathbb{N}$
  ordered by $\geq$, so that $0$ is its maximum.  Let $\X$ be the
  category associated to the order
  $(S, \leq)\oplus (\mathbb{N}, \geq)$, which by \Cref{rem:iso} is
  $\mathsf{Iso}$-adhesive. Notice that in this category the arrows
  $(a,b)\colon a\to b$ $(a,c)\colon a\to c$ do not have a pushout
  because $b$ and $c$ do not have a supremum.  Consider a rewriting
  system with set of rules $\R$ containing the following two rules
  \[\xymatrix{a & a \ar[r]^{(a, 0)} \ar[l]_{\id{a}} & 0 & a & a
      \ar[r]^{(a, b)} \ar[l]_{\id{a}} & b }\]
  We can then consider the  derivation
  $\der{D}=\{\dder{D}_i\}_{i=0}^1$ of length $2$. 
  \[
    \xymatrix@C=30pt{a \ar[d]_{m_0}&& a \ar[d]_{k_0}\ar[ll]_{\id{a}}
      \ar[r]^{(a,0)} & 0 \ar@/^.35cm/[drrr]|(.315)\hole_(.45){\id{0}}
      \ar[dr]|(.3)\hole_{\id{0}} && a \ar@/_.35cm/[dlll]^(.45){(a,c)}
      \ar[dl]|(.3)\hole^{(a,0)}& a \ar[d]^{(a,0)}\ar@{>->}[l]_{\id{a}}
      \ar[rr]^{(a,b)} && b \ar[d]^{(b,0)}\\c &&
      \ar@{>->}[ll]^{(\id{c})} c \ar[rr]_{(c,0)}&& 0 &&
      \ar@{>->}[ll]^{\id{0}} 0 \ar[rr]_{\id{0}}&& 0}\]

  \noindent
  \parbox{10cm}{ \hspace{15pt} Notice that the two
    direct derivations are sequential independent.
    %
    On the other hand, the rule
    applied by $\dder{D}_1$ cannot be applied to $c$. In fact, the
    unique morphism $a\to c$ is $(a,c)$, thus we get a diagram
    aside. But $b$ and $c$ do not have a supremum in the poset
    underlying $\X$, thus we do not get a direct derivation from
    $c$.}  \parbox{3cm}{
    \vspace{-.3cm}$\xymatrix{a \ar[d]_{(a,c)}& a
      \ar[d]_{(a,c)}\ar@{>->}[l]_{\id{a}} \ar[r]^{(a,b)} & b \\c
      & c \ar[l]^{\id{c}} }$}  
\end{example}

\iffalse

@@

  The category $\cat{Graph}$ is tame by the result in ...  Thus,
  sequential independence ensure the existence of a switch $\der{E}$,
  depicted in Fig.~\ref{fi:derE}.
  \begin{figure}
    \input{Figure/derE.tex}
    \caption{The derivation $\der{E}$.}
    \label{fi:derE}
  \end{figure}
\end{example}
\fi

@@

\todo{P: la chiamiamo ``strong indepedence pair''? o ``filler pair''}

\begin{definition}[Good independence pair]
  \label{def:filler}
  Let with $\X$ an $\mathcal{M}$-adhesive category 
  and $(\X, \R)$ a left-linear DPO rewriting system. 
  Let also $(i_0, i_1)$ be an independence pair between two direct
  derivations $\dder{D}_0$, $\dder{D}_1$
  as in the solid part of the diagram below
  \[
    \xymatrix@C=18pt{L_0 \ar[d]_{m_0}&& K_0
      \ar[d]_{k_0}\ar@{>->}[ll]_{l_0} \ar[r]^{r_0} \ar@{->}@/^.20cm/[ddrr]|(.32)\hole|(.57)\hole_(.7){u_0} & R_0
      \ar@/^.35cm/[drrr]|(.3)\hole^(.2){i_0} \ar[dr]|(.3)\hole_{h_0}
      && L_1 \ar@/_.35cm/[dlll]_(.2){i_1} \ar[dl]|(.3)\hole^{m_1}& K_1
      \ar[d]^{k_1}\ar@{>->}[l]_{l_1} \ar[rr]^{r_1} \ar@{->}@/_.20cm/[ddll]|(.32)\hole|(.57)\hole^(.7){u_1} && R_1 \ar[d]^{h_1} \\
      %
      G_0 &&
      \ar@{>->}[ll]^{f_0} D_0 \ar[rr]_(.3){g_0}&& G_1 &&
      \ar@{>->}[ll]^(.3){f_1} D_1 \ar[rr]_{g_1}&& G_2\\
      %
      & & & & P \ar@/^.3cm/@{>.>}[ull]^{p_0} \ar@/_.3cm/@{.>}[urr]_{p_1}
    }
  \]

  Consider the pullback of $g_0$ and $f_1$, getting $p_i : P \to D_i$
  ($i \in \{0,1\}$) and the mediating arrows $u_0\colon K_0\to P$ and
  $u_1\colon K_1\to P$ into the pullback given by \Cref{prop:tec} .
  %
  We will say that $(i_0, i_1)$ is a \emph{good independence pair} if
  the first two squares depicted  
  below are pushouts and if the pushout of $r_1 : K_1 \to R_1$ and $u_1 : K_1 \to P$ exists.
  \[
    \xymatrix{
      K_0 \ar[r]^{r_0} \ar[d]_{u_0}& R_0 \ar[d]^{i_0} & K_1
      \ar@{>->}[r]^{l_1} \ar[d]_{u_1}& L_1 \ar[d]^{i_1}
       &K_1 \ar[r]^{r_1} \ar[d]_{u_1}& R_1\ar@{.>}[d]^{j_0}
      \\
      %
      P \ar[r]_{p_1} & D_1 &P \ar[r]_{p_0} & D_0
       & P \ar@{.>}[r]_{q_1} & Q_1
    }
  \]
\end{definition}

@@

\subsection{Very tame rewriting systems}\label{subsec:verytame}

@@

This section is devoted to provide further details and proof for the concepts introduced in \Cref{sec:equi}.

\subsubsection{Proofs for \Cref{subsec:switch}}

@@

\subsection{Proofs for \Cref{subsec:perm}}
